Representation Learning for Generalized Planning - SAT Encodings

### 2021-04-19 doc/paper/paper.tex
encodings updated

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in
% The file ijcai21.sty is NOT the same than previous years'
\usepackage{ijcai21}

% Use the postscript times font!
\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{stmaryrd}
\urlstyle{same}

% our packages
\usepackage{amssymb}
\usepackage{environ}
\usepackage{tabularx}
% \usepackage{todonotes}
% \presetkeys{todonotes}{inline}{}
\usepackage{pgfplotstable}
\usepackage{amsmath}
\usepackage{enumitem}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\newtheorem{example}{Example}

\usepackage{comment}

% Some useful macros
%\newcommand{\rednote}[1]{\footnote{\color{red}{*** #1 ***}}}
\newcommand{\inlinecite}[1]{\citeauthor{#1}~(\citeyear{#1})}
%\newcommand{\citealp}[1]{\citeauthor{#1}~\citeyear{#1}}

%\newcommand{\predicates}[1]{\ensuremath{P(#1)}}
\newcommand{\atoms}{\ensuremath{\mathcal{H}}}
\newcommand{\actions}[1]{\ensuremath{A(#1)}}
\newcommand{\objects}[1]{\ensuremath{O(#1)}}
\newcommand{\initial}[1]{\ensuremath{I(#1)}}
\newcommand{\goal}[1]{\ensuremath{G(#1)}}
%\newcommand{\states}[1]{\ensuremath{S(#1)}}
\newcommand{\free}[1]{\ensuremath{\mathrm{free}(#1)}}
\newcommand{\vars}{\ensuremath{\mathrm{vars}}}
\newcommand{\pre}{\ensuremath{\textit{pre}}}
\newcommand{\add}{\ensuremath{\textit{add}}}
\newcommand{\del}{\ensuremath{\textit{del}}}
\newcommand{\effs}{\ensuremath{\mathrm{effs}}}
\newcommand{\states}{\ensuremath{\mathcal{S}}}
\newcommand{\transitions}{\ensuremath{\mathcal{T}}}
\newcommand{\dl}{\ensuremath{\mathcal{DL}}}
\newcommand{\labels}{L}
\newcommand{\goals}{S_G}
\newcommand{\init}{s_I}
\newcommand{\constants}{{\mathcal C}}
\newcommand{\predicates}{{\mathcal P}}
\newcommand{\Q}{\ensuremath{\mathcal Q}}
\newcommand{\successors}{\textit{succ}}
\newcommand{\dlmodel}{\mathcal M}
\newcommand{\featurespace}{\mathcal{F}}
\newcommand{\featurepool}{\ensuremath{F}}
\newcommand{\complexity}{\mathcal K}
\newcommand{\mli}[1]{\ensuremath{\textit{#1}}}
\newcommand{\dist}{\textit{dist}}
\newcommand{\training}{\ensuremath{\mathbb{T}}}
\newcommand{\apply}[1]{\ensuremath{\llbracket #1\rrbracket}}
\newcommand{\msat}{Max-SAT}

\usepackage{ textcomp }

%\hypersetup{draft}

@@

\newcommand{\pplus}{\raisebox{.15ex}{\footnotesize$\uparrow$}}
\newcommand{\mminus}{\raisebox{.15ex}{\footnotesize$\downarrow$}}
\newcommand{\nils}{\raisebox{.15ex}{$\textdegree$}}

% \newcommand{\wip}[1]{{\color{red} #1}}  % From "work in progress" :-)
% \newcommand{\gfm}[1]{\footnote{\color{red}{[Guillem] #1}}}
% \newcommand{\F}{\ensuremath{\mathcal{F}}}
% \renewcommand{\S}{\mathcal{S}}
% \newcommand{\G}{\mathcal{G}}

% the following package is optional:
%\usepackage{latexsym}

% Following comment is from ijcai97-submit.tex:
% The preparation of these files was supported by Schlumberger Palo Alto
% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
% preparation.

% These instructions can be modified and used in other conferences as long
% as credit to the authors and supporting agencies is retained, this notice
% is not changed, and further modification or reuse is not restricted.
% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
% contacts for providing assistance without their prior permission

% To use for other conferences, change references to files and the
% conference appropriate and use other authors, contacts, publishers, and
% organizations.
% Also change the deadline and address for returning papers and the length and
% page charge instructions.
% Put where the files are available in the appropriate places.

%PDF Info Is REQUIRED.
\pdfinfo{
/TemplateVersion (IJCAI.2021.0)
}

\title{Learning Generalized Policies Over Grid Environments}

% \title{General Polizy Learner for Grid Environments}

\author{

}

% Papers must be no longer than 7 pages in total: 6 pages for the body of the paper (including all figures/tables),
% plus up to 1 additional page with references
\begin{document}

@@

\begin{abstract}
\end{abstract}

\section{Introduction}

\section{Logic Language}

The Grid Domains share the set of Objects \textit{O}, and the set of predicates \textit{P}.

\subsection{Objects}

All objects in \textit{O} belong to one sort, \textit{cell}. Then, we have an object \textit{o} for each cell in the grid. For instance, in a grid of size $i \times j$ the set \textit{O} contains the object \textit{cij}, representing the cell in the row \textit{i} and the column \textit{j}.

\subsection{Predicates}

To explain the predicates it is good to remember that each of the domains has different values contained in the grid, e.g. the domain reach-for-the-star has five possible values for each cell in the grid, which are: Empty, drawn, agent, star, right-arrow, and left-arrow. Then, let us define that for each domain there is a set of values \textit{V} which represents the values that each cell object in the grid can take. It is also good to remember the difference between static and dynamic predicates. Then, the static predicates are always the same for any state of an instance, while a dynamic predicate will change depending on the state of the instance. \textit{P} will contain both types of predicates, dynamic and static.\\

We have used the dynimic predicates to define the value of \textit{V} that any object \textit{o} can take at a given time. While we have used the static predicates to define the topology of the grid, i.e., the relative position of o with respect to o'. It is important to note that it is always preferable to have a static predicate rather than a dynamic one as this will considerably reduce the problem of the combinatorial explosion of symbols [some reference]. Finally, we define a set of relative directions \textit{D}, which contains \{right, left, down, up, right-up, right-down, left-down, left-up, horizontal, vertical\}. Then, \textit{D} is used to define the topology of the grid.\\

Then, \textit{P} contains:

\subsubsection{Dynamic predicates:}

\begin{itemize}

\item \textit{v(cij)}: Means that cell \textit{o} has the value \textit{v} at a certain time. Recall that v is a concrete value within the set \textit{V}. For instance, if \textit{v} is equalt to the value \textit{agent}, then \textit{agent(cij)} means that cell \textit{cij} has the value agent.
\item \textit{player(n)}: This is predicate used for multiplayer games where $n \in \mathbb N$, and represents the player whose turn it is this turn. For instance, the predicate \textit{player(0)} means that in a specific state is the turn of the player \textit{0}.

\end{itemize}

\subsubsection{Static predicates:}

\begin{itemize}

\item \textit{d(o, o')}: means that \textit{o} has \textit{o'} as contiguous in direction d. For example: \textit{right(cij, cij')} means that to the adjacent right of cell \textit{cij} is the cell \textit{cij'}. Notice that some directions of \textit{D} are redundan, e.g., the predicate \textit{horizontal(cij, cij')} means that to the adjacent right o left of cell \textit{cij} is the cell \textit{cij'}, and could be defined combining two of the other predicates like $\textit{right(cij, cij')} \lor \textit{left(cij, cij')}$. In a similar way \textit{vertical(cij, cij')} can be defined as $\textit{up(cij, cij')} \lor \textit{down(cij, cij')}$. Furthermore, some directions like \textit{left(cij, cij')} are also redundant, since can be redefined as the invers of its opposite direction like \textit{right(cij', cij)}. This redundancy will increase the number of possible features according to the grammar, however it will also reduce the complexity of the features.

\item \textit{same-row(o, o')}: Means that \textit{o} is in the same row as \textit{o'}. For example: \textit{same-row(cij, cij')} means that cell \textit{cij} is in the same row as cell \textit{cij'}, the row \textit{i}.

\item \textit{same-col(o, o')}: Means that \textit{o} is in the same column as \textit{o'}. For example: \textit{same-col(cij, ci'j)} means that cell \textit{cij} is in the same column as cell \textit{ci'j}, the column \textit{j}.

\end{itemize}

\subsection{Encodings}

\subsubsection{FOND Adversarial}

\subsubsection{Main Variables}
\begin{itemize}
 \item $Good(s, s')$ for $s$ alive, $s'$ solvable and $(s, s') \not\in \badtx$.

@@

@@

% ... End

%\section*{Acknowledgments}

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\small
\bibliographystyle{named}
\bibliography{abbrv-short,literatur,references,crossref-short}

\end{document}

>>>>

\documentclass[a4paper]{article}
%\def\year{2020}\relax
%\usepackage{aaai20}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\frenchspacing  % DO NOT CHANGE THIS
%\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
%\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS

\usepackage[margin=1in]{geometry}

% our packages
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{bm}
\usepackage{booktabs}
%\usepackage[inline]{enumitem}
%\usepackage{mathtools}
%\usepackage{multirow}
%\usepackage[mode=buildnew,subpreambles=true]{standalone}
\usepackage{subcaption}
%\usepackage{todonotes}
\usepackage[table]{xcolor}  % TODO Comment out for final submission
%\usepackage{pgfplots}
%\usepackage{tikz}
\usepackage{natbib}
\usepackage{fancyvrb}

% Some useful macros
%\newcommand{\inlinecite}[1]{\citeauthor{#1}~(\citeyear{#1})}
%\newcommand{\citealp}[1]{\citeauthor{#1}~\citeyear{#1}}

\newcommand{\smallpar}[1]{{\vspace{10pt}\noindent \bf #1.}}

\newcommand{\free}[1]{\ensuremath{\mathrm{free}(#1)}}
\newcommand{\vars}{\ensuremath{\mathrm{vars}}}
\newcommand{\pre}{\ensuremath{\mathrm{pre}}}
\newcommand{\add}{\ensuremath{\mathrm{add}}}
\newcommand{\del}{\ensuremath{\mathrm{del}}}
\newcommand{\effs}{\ensuremath{\mathrm{effs}}}

@@

\newtheorem{definition}{Definition}
\newtheorem{definitionandtheorem}[definition]{Definition and Theorem}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{corollary}[definition]{Corollary}
\newtheorem{example}[definition]{Example}

\newcommand{\wip}[1]{{\color{red} #1}}  % From "work in progress" :-)
\newcommand{\gfm}[1]{\footnote{\color{red}{[Guillem] #1}}}

\newcommand{\numtasks}[1]{\small{(#1)}}

\newcommand{\badtx}{\ensuremath{\mathrm{BAD}}}

\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.

%PDF Info Is REQUIRED.
% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
 \pdfinfo{
/Title ()
/Author ()
} %Leave this

\title{Representation Learning for Generalized Planning - SAT Encodings}
%Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
%\author{GFM}

\begin{document}

@@

\section{Learning of Generalized Features \& Policy in Adversarial FOND domains}

Encoding is parametrized by
\begin{itemize}
 \item Pool $F$ of description logic features $f$, each with given feature complexity $\mathcal{K}(f)$.
 \item Training set consisting of a sample of transitions from a number of instances of the same domain.
       At the moment we're assuming the sample is complete, and we have full information on whether each state
       in the sample is a goal state, an unsolvable state, or otherwise (see below for definitions). We also have access
       to the minimum distance to a goal $V^*(s)$ for each state $s$.
 \item A parameter $\delta$ which is a ``slack'' value to determine the maximum deviation from the optimal $V^*(s)$
 what we will allow in our policy. This will be made clearer in the encoding below.
\end{itemize}

\subsection{Terminology}
A state is called \emph{reachable} if there is a path to it from $s_0$, and
it is called \emph{solvable} if there is a path from it to a goal
state.
A state is \emph{alive} if it is solvable, reachable and not a
goal state~\cite{frances-et-al-ijcai2019}.

We use $T$ to denote the set of all transitions $(s, s')$ in the training sample such that $s$ is alive.

\subsection{Improvements}

\paragraph{Non-distinguishability of transitions as an equivalence relation.}
Any fixed, given pool of features $F$ implicitly defines an equivalence relation where to transitions are
equivalent iff they cannot be distinguished \emph{by any feature in $F$}.
If two transitions cannot be distinguished by any feature, then clearly either the policy computed by the SAT solver
considers all of them as ``good'', or as ``bad''.
We'll exploit this by using one single SAT variable to denote whether \emph{all transitions in a given equivalence
class} are good or bad. When exploiting this notion of equivalence (which is implemented as an optional feature of
the CNF generator), then every mention below to SAT variable $Good(s, s')$ needs to be read as $Good(s_{\star}, s_{\star}')$,
where $(s_{\star}, s_{\star}')$ is the \emph{representative} transition of the equivalence class to which $(s, s')$ belongs.

\paragraph{``Bad'' transitions.}
We use \badtx{} to denote the set of transitions that have been determined at preprocessing as necessarily
\emph{not} good.
At the moment, this set contains all transitions that go from an alive to an unsolvable state and, if using
equivalence relations, all those other transitions in whose equivalence class there is some other ``bad'' transition.

\paragraph{Feature dominance.}
Work In Progress.\footnote{This is implemented, but needs to be adapted to the equivalence classes improvement.}

\newpage

\subsection{Variables}

\subsubsection{Main Variables}

Let the transitions be defined as $s \rightarrow a \rightarrow s' \rightarrow s"$, where the agent executes a in the state s and goes to some $s'$, and then the nature or the adversary (doing anything) turns $s'$ into $s''$. Therefore the transitions $(s, s')$ is deterministic and the transition $(s, s'')$ is non-deternimistic.

\begin{itemize}
 \item $Good(s, a)$ for $s$ alive, $s''$ solvable. Where $V(s'') < V(s)$.

 \item $Good(s, s')$ for $s$ alive, $s'$ solvable and $(s, s') \not\in \badtx$.

@@

\smallpar{C3 (Optional)} All descending transitions must be considered Good:
\begin{align}
 V(s, d) \land V(s', d') \rightarrow Good(s, s'),&\;\; \text{for $s, s'$ alive, $(s, s') \not\in \badtx$, $1 \leq d' < d \leq D$,} \\
 Good(s, s'),&\;\; \text{for $s$ alive, $s'$ goal.} \tag{\theequation${}^\prime$}
\end{align}

@@

\section{Empirical Results}

\bibliographystyle{plain}
% Cross-referenced entries need to go after the entries that cross-reference them
\bibliography{abbrv-short,literatur,references,crossref-short}
\end{document}
